The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-04-10 paper/iphas-dr2.tex

A new catalogue of 219 million sources has been
derived from the INT/WFC Photometric \ha\ Survey
of the Northern Galactic Plane.
It is the first catalogue to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
taking in the Northern Milky Way at $|b|<5^\circ$.
The new 97-column catalogue provides single-epoch photometry
across 92\% of the survey area,
and is the first quality-controlled and uniformly calibrated
catalogue to have been constructed from the imaging data.
We explained the data reduction and quality control procedures that
were applied,
described and tested the global photometric calibration,
and detailed the construction of the source catalogue.

The observations included in this release
show a median seeing of 1.1 arcsec
and $5\sigma-$depths of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$,
and \ha$=20.3\pm 0.3$.
The global calibration and photometric repeatability
were found to be accurate at the level of $0.03$ mag (rms),
providing a significant improvement over the 
previous data release.
The source catalogue provides the best-available
single-epoch astrometry and photometry
for 219~million unique sources.
We detailed a list of recommended quality criteria
to select objects with reliable colours from the catalogue.
Finally, we demonstrated the use of the survey's unique
$(r-\ha,\ r-i)$ diagram for characterising stellar populations
and selecting emission-line objects.

In future work, we intend to draw upon the PanSTARRS photometric
survey to further improve the accuracy of our global calibration.
We also plan to produce a catalogue 
which details all the secondary detections
to aid time-domain studies,
and are investigating the use of photometry
derived from PSF-fitting across the survey.

The data-taking strategy developed for IPHAS
has since been reapplied to carry out 
a companion INT/WFC Galactic Plane survey called UVEX
in U, $g$, $r$, He {\sc i} \citep{Groot2009},
a survey of the Kepler field 
in $U$, $g$, $r$, $i$, \ha\ 
\citep{Greiss2012},
and a survey of the entire Southern Galactic Plane
in $u$, $g$, $r$, $i$, \ha\ 
called VPHAS+ \citep{Drew2014}.
We intend to use the present work
as a template for generating catalogues
from these companion surveys.
When combined, these surveys
will offer new insights
into the contents and structure of the Milky Way
across 360$^\circ$,
and provide images across the most crowded
regions to be faced by Gaia.

>>>>

A new catalogue of 219 million sources has been
derived from the INT/WFC Photometric \ha\ Survey
of the Northern Galactic Plane.
It is the first catalogue to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
taking in the Northern Milky Way at $|b|<5^\circ$.
The new 97-column catalogue provides single-epoch photometry
across 92\% of the survey area,
and is the first quality-controlled and uniformly calibrated
catalogue to have been constructed from the imaging data.
%We explain the data reduction and quality control procedures that
%were applied.  We describe and test the global photometric 
%calibration, 
%and detailed the construction of the source catalogue.

The observations included in this release
achieve a median seeing of 1.1 arcsec
and $5\sigma-$depths of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, and \ha$=20.3\pm 0.3$.
The global calibration and photometric repeatability
are found to be accurate at the level of $0.03$ mag (rms),
providing a significant improvement over the 
previous data release.
The source catalogue specifies the best-available
single-epoch astrometry and photometry
for 219~million unique sources.
To support its exploitation, we provide a list of recommended quality criteria
that will permit the selection of objects with reliable colours from 
the catalogue.  The closing demonstrations highlight the use of the 
survey's unique $(r-\ha,\ r-i)$ diagram for characterising stellar populations
and selecting emission-line objects.  More comprehensive applications of IPHAS
can be found in the works of Sale et al (2014), which applies DR2 to the problem
of 3D extinction mapping, and of Sabin et al (2014), where the results of a search
of the image database for new planetary nebulae is presented.

The current plan is to work toward one further major IPHAS source catalogue, in which 
the remaining gaps in sky coverage will have been eliminated -- observations aimed at 
replacing data not meeting the IPHAS minimum quality requirements continue.  We will also 
examine options to further improve the global calibration, perhaps tightening the accuracy to
better than 2\%.  For example, we will investigate the use of e.g. the PanSTARRS photometric 
ladder (Magnier et al 2013) as a reference set, and explore improving source recovery in 
the most dense fields via the implementation of PSF fitting in place of aperture photometry.  
Finally, the next catalogue will detail all the secondary detections to aid time-domain studies.

The data-taking strategy developed for IPHAS
has since been reapplied to carry out 
a companion INT/WFC Galactic Plane survey called UVEX
in U, $g$, $r$, He {\sc i} \citep{Groot2009},
a survey of the Kepler field 
in $U$, $g$, $r$, $i$, \ha\ 
\citep{Greiss2012},
and a survey of the Southern Galactic Plane and Bulge
in $u$, $g$, $r$, $i$, \ha\ 
called VPHAS+ \citep{Drew2014}.  The work presented here
stands as a potential template for the catalogues that
remain to be generated from these sibling surveys.  
In prospect from them, whether they are mined separately or together,
are the means to ask seamless questions on the contents and structure of 
the most highly-populated components of the Milky Way.
